---
title: "Peer Assessment 1"
author: "Jurijs Jefimovs"
date: "Tuesday, July 15, 2014"
output: html_document
---
```{r LOAD_LIBRARIES }
require(ggplot2)
```

###Loading and preprocessing the data
1.Load the data
```{r DataLoad}
unzip("activity.zip")
data <- read.csv("activity.csv")
str(data)
```

2. Data transformation. Converting date column from factor to Date type.
```{r DataTransformation}
data$dt <- as.Date(data$date, format="%Y-%m-%d")
str(data)
```

###What is mean total number of steps taken per day?
1. Histogram of the total number of steps taken each day
```{r Histogram}
hist_data <- aggregate(steps ~ date, data, FUN = sum)
hist(hist_data$steps, main = "Histogram of the total number of steps taken each day", xlab = "Steps")
```

2. The mean and median total number of steps taken per day
```{r MeanAndMedian}
summary(hist_data$steps)
summary(hist_data$steps)[c("Mean","Median")]
```

###What is the average daily activity pattern?
1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r TimeSeriesplot}
tseries_by_interval <- aggregate(steps ~ interval, data, FUN = mean)
plot(tseries_by_interval$interval, tseries_by_interval$steps, type = "l", main = "Average number of steps across all days", xlab = "Interval", ylab = "Average number of steps")
```

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r MaxNumberOfSteps}
tseries_by_interval$interval[tseries_by_interval$steps == max(tseries_by_interval$steps)]
```

###Imputing missing values

1. The total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r TotalNumberOfNa}
sum(is.na(data$steps))
```

2. Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

Answer: The strategy for filling missing values is to use median for 5-minute interval across all days.
```{r MissingValuesStrategy}
median_by_interval <- aggregate(steps ~ interval, data, FUN = median)
```

3. Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r NewDataSet}
rownames(median_by_interval) <- median_by_interval$interval
na_steps <- data$interval[is.na(data$steps)]
data$steps[is.na(data$steps)] <- median_by_interval[as.character(na_steps),"steps"]
```

4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day.
```{r HistogramAfterFillingNaValues}
hist_data <- aggregate(steps ~ date, data, FUN = sum)
hist(hist_data$steps, main = "Histogram of the total number of steps taken each day", xlab = "Steps")

summary(hist_data$steps)
summary(hist_data$steps)[c("Mean","Median")]

```

Question 1:     Do these values differ from the estimates from the first part of the assignment?
Answer 1:       Yes, they do. After filling missing values, Mean and Median is shifted to the left.

Question 2:     What is the impact of imputing missing data on the estimates of the total daily number of steps?
Answer 2:       It is obvios that after imputing data, totals has increased where data was imputed.

###Are there differences in activity patterns between weekdays and weekends?
1. Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

* NOTE: weekday() function prints weekdays using local regional settings. (for instance if you have russian locale, you will get russians weekday names).Instead i use format() with "format" = %u (Weekday as a decimal number 1–7, Monday is 1)

```{r }
data$weekday <- format(data$dt, format = "%u") #see *NOTE above
data$weekday_name[data$weekday < 6] <- "weekday"
data$weekday_name[data$weekday >= 6] <- "weekend"
```

2. Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).

```{r }
tseries_by_weekday_interval <- aggregate(steps ~ interval + weekday_name, data, FUN = mean)
qplot(interval, steps, data=tseries_by_weekday_interval, geom="line",facets = weekday_name~. )

```
